---
title: "Reproducible research -- Project 1"
author: "Gururaj"
date: "5/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview  
As part of this assignment, analysing the data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up.

### Load relevant libraries

```{r loadLibs}
library("data.table")
library(ggplot2)
```
### Prepare data for analysis
```{r getData}
activitydata = read.csv("activity.csv", header=TRUE, sep=",")
str(activitydata)
```
### Format activity data
```{r formatData}
activitydata$date = as.Date(activitydata$date, format="%Y-%m-%d")
activitydata$interval = as.factor(activitydata$interval)
str(activitydata)
head(activitydata,10)
```
### 1. Total number of steps taken per day
```{r calc1}
totalStepsPerDay = aggregate(steps ~ date, data=activitydata, FUN=sum)
colnames(totalStepsPerDay) = c("date", "steps")
head(totalStepsPerDay,10)
```
### 2. Histogram of the total number of steps taken each day
```{r plot1}
g = ggplot(totalStepsPerDay, aes(x = steps)) +
  geom_histogram(color="black", fill="grey", binwidth = 1000) +
  labs(title = "Histogram for Steps per Day", x = "Steps", y = "Frequency")
print(g)
```
### 3. Mean and median for number of steps taken each day  

```{r calcMean}
meanOfStepsPerDay =  mean(totalStepsPerDay$steps)
paste("Mean of Steps per Day: " , meanOfStepsPerDay)
mediaOfStepsPerDay = median(totalStepsPerDay$steps)
paste("Median of Steps per Day: " , mediaOfStepsPerDay)
```

### 4. Time series plot of the average number of steps taken

```{r calc2}

stepsPerInterval = aggregate(steps ~ interval, data = activitydata, FUN = mean, na.rm = TRUE)
stepsPerInterval$interval = as.integer(levels(stepsPerInterval$interval)[stepsPerInterval$interval])
colnames(stepsPerInterval) = c("interval", "steps")
head(stepsPerInterval,5)

```
```{r plot2}
g = ggplot(stepsPerInterval, aes(x = interval, y = steps)) + 
            geom_line(col = "blue", size = 1)+
            labs(title = "Average Daily Activity-- Pattern", x = "Interval", y = "Steps")

print(g)

```  

### 5. The 5-minute interval that, on average, contains the maximum number of steps  

```{r calc3}
maxInterval= stepsPerInterval[which.max(stepsPerInterval$steps),]
maxInterval
```

## Imputing missing values

### 1 NUmber of missing values in the activity data set
```{r calc4}

missingValues = sum(is.na(activitydata$steps))
paste(" Missing values in the activity data set are: " , missingValues)
```

### 2. Strategy for filling in all of the missing values in the dataset.  

For populating missing values, suggest to replace missing values with the mean values at same intervals accross days!

### 3. New data set by replacing missing values
```{r calc5}
newData = activitydata
indexOfNA = which(is.na(newData$steps))
  for (i in indexOfNA) 
    {
      newData$steps[i] <- with(stepsPerInterval, steps[interval = newData$interval[i]])
  }
head(newData,10)

# Check if there are any missing values in the new data set
missingValues = sum(is.na(newData$steps))
paste(" Missing values in the new activity data set are: " , missingValues)

```
### 4. Analyse New data after replacing the missing values  

Analysing the new data set by looking at the histogram, mean/median values below:

### 4.1. Total number of steps taken per day
```{r calc11}
totalStepsPerDay_1 = aggregate(steps ~ date, data=newData, FUN=sum)
colnames(totalStepsPerDay_1) = c("date", "steps")
head(totalStepsPerDay_1,10)
```
### 4.2. Histogram of the total number of steps taken each day
```{r plot11}
g1 = ggplot(totalStepsPerDay_1, aes(x = steps)) +
  geom_histogram(color="black", fill="blue", binwidth = 1000) +
  labs(title = "Histogram for Steps per Day(without NAs)", x = "Steps", y = "Frequency")
print(g1)
```

### 4.3. Mean and median for number of steps taken each day  

```{r calcMean1}
meanOfStepsPerDay_1 =  mean(totalStepsPerDay_1$steps)
paste("Mean of Steps per Day: " , meanOfStepsPerDay_1)
mediaOfStepsPerDay_1 = median(totalStepsPerDay_1$steps)
paste("Median of Steps per Day: " , mediaOfStepsPerDay_1)

```
```{r calcMean12,echo=FALSE}

paste("Old Mean: " ,meanOfStepsPerDay, "Old Median: " , mediaOfStepsPerDay, 
                  "Difference is", (meanOfStepsPerDay-mediaOfStepsPerDay) )
paste("New Mean: " ,meanOfStepsPerDay_1, "New Median: " , mediaOfStepsPerDay_1,
                  "Difference is", (meanOfStepsPerDay_1-mediaOfStepsPerDay_1))

```
After replacing the missing values, there is no difference between mean and median values!

## Review of activity patterns between weekdays and weekends  

Adding factor variable to denote a particular day is a weekday or weekend

```{r calc7}

data1 = data.table(newData)
data1[, weekday := ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")]
data1$weekday <- as.factor(data1$weekday)
data1$interval <- as.integer(levels(data1$interval)[data1$interval])
head(data1, 5)
tail(data1, 5)
```
### Time series plot

```{r plot8}
stepsPerWeekday = aggregate(steps ~ interval+weekday, data = data1, FUN = mean)
ggplot(stepsPerWeekday, aes(x = interval, y = steps)) + 
  geom_line(col = "red", size = 1) + 
  facet_wrap(~ weekday, nrow = 2, ncol = 1) + 
  labs(title = "Average Daily Activity-- Pattern", x = "Interval", y = "Steps")

```





