---
title: "Spend Analysis using R & Shiny App"
author: "Gururaj"
date: "8/6/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

As part of this assignment, reviewed the expenses of an individual in the year 2019 using R and Shiny App.  
Following are the inputs for the analysis:  
1. CSV data @  https://jguru2020.github.io/data/Expenses.csv  
2  Libraries like **dplyr** & **Plotly**  
3. R & Shiny App  
4. Shinyapps.io

```{r load, echo = FALSE, message=FALSE}
library(dplyr)
library(plotly)
spendData = read.csv("https://jguru2020.github.io/data/Expenses.csv")
```

## Spend analysis by Histogram

```{r plot-1, echo = FALSE, message=FALSE}

plot1 <- plot_ly(spendData, x = ~category) %>% add_histogram()

plot1

```  

## Spend analysis by Pie Chart

```{r plot-2, echo = FALSE, message=FALSE}

plot2 <- plot_ly(spendData, labels = ~category, values = ~spend, type = 'pie')
plot2 <- plot2 %>% layout(title = 'Spend Analysis',
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

plot2
```


## Challenges  

Above shown plots are stand-alone / static plots. When we have more sub catories for each category, 
we have to create those many plots.   

Challenges:  
- Drill-down of the plots is not possible  
- Unable to access analysis from web/mobile/tablet  

## Solution  

Using Shiny App is the solution. With Shiny app we could:

- Access the analysis from anywhere in the world (web/mobile/tablet)
- Look & Feel is same accross the platforms (web/mobile/tablet)
- Allows user to click on a slice of the pie to ‘drill-down’ into sub-categories of the chosen category  

Souce code @ https://github.com/jguru2020/jguru2020.github.io/tree/master/dataproducts/week4

## Demo

Access Shiny App @ https://jguru2020.shinyapps.io/spend_analysis/

- Click on category like **Merchandise**
- See all the sub-categories of Merchandise
- Click **Back** button to see pie chart with main categories
- Click on category like **Transportaion**
- See all the sub-categories of Merchandise
- Click **Back** button to see pie chart with main categories
- Continue...  



## Thank you!!


